---
title: "World Happiness "
author: "Julian Usuga Ortiz"
date: "24/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = FALSE)
library(tidyverse)
library(gganimate)
library(gifski)
library(png)
```


## This is an exploratory analysis to world happiness scores
<br><br><br><br>

#### See info and data source 

<https://www.kaggle.com/unsdsn/world-happiness>

<br><br>
First i begin cleaning and transforming my data


```{r}
data15 <- read.csv(file = '2015.csv')
data15$year = "2015"
data16 <- read.csv(file = '2016.csv')
data16$year = "2016"
data17 <- read.csv(file = '2017.csv')
data17$year = "2017"
data18 <- read.csv(file = '2018.csv')
data18$year = "2018"
data19 <- read.csv(file = '2019.csv')
data19$year = "2019"


data15 <- data15 %>% select(!Dystopia.Residual & !Standard.Error & !Family &!Region)

data16 <- data16 %>% select(!Dystopia.Residual & !Upper.Confidence.Interval & !Lower.Confidence.Interval & !Family &!Region)

data17 <- data17 %>% select(!Whisker.high & !Whisker.low & !Dystopia.Residual & !Family)

data18 <- data18 %>% select(!Social.support)

data19 <- data19 %>% select(!Social.support)


data15 <- rename(data15,
                 c('Country.or.region'='Country'), 
                 c("Overall.rank"="Happiness.Rank"), 
                 c("GDP.per.capita"="Economy..GDP.per.Capita."), 
                 c("Freedom.to.make.life.choices"="Freedom"), 
                 c("Perceptions.of.corruption"="Trust..Government.Corruption."), 
                 c("Score"="Happiness.Score"), 
                 c("Healthy.life.expectancy"="Health..Life.Expectancy."))
data16 <- rename(data16,
                 c('Country.or.region'='Country'), 
                 c("Overall.rank"="Happiness.Rank"), 
                 c("GDP.per.capita"="Economy..GDP.per.Capita."), 
                 c("Freedom.to.make.life.choices"="Freedom"), 
                 c("Perceptions.of.corruption"="Trust..Government.Corruption."), 
                 c("Score"="Happiness.Score"), 
                 c("Healthy.life.expectancy"="Health..Life.Expectancy."))
data17 <- rename(data17,
                 c('Country.or.region'='Country'), 
                 c("Overall.rank"="Happiness.Rank"), 
                 c("GDP.per.capita"="Economy..GDP.per.Capita."), 
                 c("Freedom.to.make.life.choices"="Freedom"), 
                 c("Perceptions.of.corruption"="Trust..Government.Corruption."), 
                 c("Score"="Happiness.Score"), 
                 c("Healthy.life.expectancy"="Health..Life.Expectancy."))

happiness <- rbind(data15, data16, data17, data18, data19)


#Remove initial data
rm(data15, data16, data17, data18, data19)
```

```{r,include=FALSE, message=FALSE}
happiness$Perceptions.of.corruption <- as.numeric(as.character(happiness$Perceptions.of.corruption) )
```

<br><br>
Here we can see the range of each variable, the higher the number, the higher the indicator


```{r}
summary(happiness)
```

<br><br>
Violin plot gives us the impression that since 2017 the 30 countries with the highest happiness index are becoming healthier


```{r, message=FALSE}
happiness %>%
   filter(Overall.rank < 30) %>%
   ggplot(aes(x =year, y = Healthy.life.expectancy, color = year)) +
   geom_violin()+
   geom_jitter() +
   ylab("Healthy life Expectancy")
```


Looks like people are becoming healthier but happiness scores remain at the same levels with the top 50 countries in the index


```{r, message=FALSE}

happiness %>%
   filter(Overall.rank < 50) %>%
   ggplot(aes(x = Healthy.life.expectancy, y = Score, color = year)) +
   geom_jitter()+
   geom_smooth(method = "lm", se = FALSE)+
   ylab("Happiness Score") +
   xlab("Healthy Life Expectancy") 
```

<br><br>
With low confidence we can say that the top 50 countries in the happiness index are happier but not free


```{r, message=FALSE}

happiness %>%
   filter(Overall.rank < 50 & (year == "2015" | year == "2019")) %>%
   ggplot(aes(x =  Freedom.to.make.life.choices, y = Score, color = year)) +
   geom_jitter()+
   geom_smooth(method = "lm")+
   ylab("Happiness Score") +
   xlab("Freedom to make life choices") 
```

<br><br>
And here's an animated graph for each year!


```{r, message=FALSE}

p <- happiness %>%
   filter(Overall.rank < 50) %>%
   ggplot(aes(x = Healthy.life.expectancy, y = Score, color = "red")) +
   geom_jitter()

anim <- p + transition_states(year,
                    transition_length = 2,
                    state_length = 1)+
   labs(title = 'Year: {closest_state}', x = 'Life expectancy', y = 'Happiness Score')
anim
```

